---
title: "Simulation Exercise"
author: "Marque Staneluis"
date: "April 28, 2018"
output:
  html_document: default
  pdf_document: default
---

## Overview
The purpose of this project is to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 
Set lambda = 0.2 for all of the simulations. This investigates the distribution of averages of 40 exponentials over 1,000 simulations.


## Part 1: Simulations
Set the simulation variables lambda, n, and seed.
```{r}
ECHO=TRUE
set.seed(1337)
lambda = 0.2
n = 40
```

Run Simulations with variables
```{r}
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(n, lambda)))
```


## Part 2: Sample Mean versus Theoretical Mean
#### Sample Mean
Calculating the mean from the simulations with give the sample mean.
```{r}
sampMean <-mean(mns)
sampMean
```

#### Theoretical Mean
The theoretical mean of an exponential distribution is lambda^-1.
```{r}
theoMean <- lambda^-1
theoMean
```

#### Comparison
There is only a slight difference between the simulations sample mean and the 
exponential distribution theoretical mean.
```{r}
abs(mean(mns)-lambda^-1)
hist(mns, main = "Sample Mean versus Theoretical Mean", col = "white", breaks = 100)
abline(v = sampMean, col = "red")
abline(v = theoMean, col = "blue")
```


## Part 3: Sample Variance versus Theoretical Variance
#### Sample Variance
Calculating the variance from the simulation means with give the sample
variance.
```{r}
sampVar <- var(mns)
sampVar
```

#### Theoretical Variance
The theoretical variance of an exponential distribution is 
(lambda * sqrt(n))^-2.
```{r}
theoVar <- (lambda * sqrt(n))^-2
theoVar
```

#### Comparison
There is only a slight difference between the simulations sample variance and 
the exponential distribution theoretical variance.
```{r}
abs(var(mns)-(lambda * sqrt(n))^-2)
```


## Part 4: Distribution
This is a density histogram of the 1,000 simulations. There is an overlay with a 
normal distribution that has a mean of lambda^-1 and standard deviation of 
(lambda*sqrt(n))^-1, the theoretical normal distribution for the simulations.
```{r}
hist(mns, main = "Normal Distribution", col = "yellow", breaks = 100)
xfit <- seq(min(mns), max(mns), length = 100)
yfit <- dnorm(xfit, mean = 1/lambda, sd = (1/lambda)/sqrt(n))
lines(xfit, yfit*60, lty = 5)
```
